---
title: "Deliverable1"
output: html_notebook
---
install packages for analysis
```{r}
installed.packages("rio")
installed.packages("ggplot2")
installed.packages("DescTools")
installed.packages("dplyr")
```

```{r}
library(rio)
install_formats()
rm(list = ls())

library(ggplot2)
library(DescTools)
library(dplyr)
```

Getting the data TABLE from the file
```{r}
link='https://github.com/superadlerrrk/543/blob/main/Deliverable1/Crime_Data.csv?raw=true'
crime = import(link)
```

This is the list of the variables (columns) from the data loaded
...and the first thirty values:
```{r}
names(crime)
head(crime$"Primary Offense Description",30) 
```

Checking which columns contain null values
```{r}
cols_with_nulls <- colSums(is.na(crime))
```

Printing the column names with null values
```{r}
names(crime)[cols_with_nulls > 0]

any(is.na(crime$"Reported Time"))
any(is.na(crime$"Occurred Time"))

sum(is.na(crime$"Reported Time"))
sum(is.na(crime$"Occurred Time"))
```

Creating a logical vector indicating which rows have complete data in my_column
```{r}
complete_rows <- complete.cases(crime$"Reported Time")
complete_rows <- complete.cases(crime$"Occurred Time")
```

Get absolute values by each Precinct
```{r}
absoluteT=table(crime$Precinct)
absoluteT
```

Transform the absolute count into percent or relative frequencies:
```{r}
prop.table(absoluteT)
```

Transform the range of the proportions into percents:
```{r}
propT=prop.table(absoluteT)*100
propT
```

Get the pie plots by Precinct:
```{r}
pie(absoluteT, cex = 0.7)
```

Subset the data frame to remove rows with null values in my_column
```{r}
crime <- crime[complete_rows, ]

sum(is.na(crime$"Reported Time"))
sum(is.na(crime$"Occurred Time"))

crime <- crime %>%
  mutate(Precinct = if_else(Precinct == "", "UNKNOWN", Precinct))

crime_grouped <- crime %>%
  group_by(Precinct, Neighborhood) %>%
  summarise(n = n())

crime_totals <- crime_grouped %>%
  group_by(Precinct) %>%
  summarise(total_entries = sum(n))

crime_grouped <- crime_grouped %>%
  left_join(crime_totals, by = "Precinct") %>%
  mutate(ratio = n / total_entries)

precincts <- unique(crime_grouped$Precinct)

for (precinct in precincts) {
  data <- filter(crime_grouped, Precinct == precinct)
  plot_title <- paste0(precinct, "Crime Entries by Neighborhood")
  
  pie_chart <- ggplot(data, aes(x="", y=ratio, fill=Neighborhood)) +
    geom_bar(stat="identity", width=1) +
    coord_polar("y", start=0) +
    labs(title=plot_title) +
    theme_void()
  
  print(pie_chart)
}
```
